\pkg{cubble}: An R Package for Structuring Spatio-temporal Data

### 2022-01-27 4-3-water-level.Rmd
clean up water level example code and a draft on its example section

The water level data comes from [Bureau of Meteorology](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) and has a copy in `weatherdata`. Here we extract the water course level and add a column annotate this data of type `river`. For the rainfall data, we will still use the `weatherdata::climate_full`, filtering for Victorian stations in 2020 should be pretty familiar by now. Again, we first look at where these stations are on the map first:

Now we use `match_sites()` to first pair the weather stations with the river gauges spatially and then apply the temporal matching on `prcp`. We will construct the interval based on peaks in `climate` since we would expect a lag effect for precipitation to flow into the river and cause a raise in river level, hence `temporal_independent = climate`. We select the 30 highest peak from the series to construct the match by setting `temporal_n_highest = 30`. This is a tuning parameter and you can start with 10% of the points of one series (here we have daily data for a year, 10% is roughly 30 points). `temporal_min_match` filters out pairs don't have enough match and to return all the pairs, set `temporal_min_match` to `0`.``

The output from temporal matching is also a cubble, with additional column `.dist` and `.group` inherent from spatial matching and `n_match` for the number of matched temporal peaks. Then you can use this output to plot the location of match or to look at the series:

```{r matching, out.width="100%", out.height="30%", fig.cap="asdfasd"}
knitr::include_graphics("figures/matching.png")
```

>>>>

Water data is collected by the [Bureau of Meteorology](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full). This includes ... water gauges, variables collected. In particular, water level will interactive with the climate variable precipitation as rains fall into the river will raise the water level. Figure \ref{fig:matching-map} shows the available weather station and water gauges in Victoria, Melbourne.

```{r echo = TRUE}
climate <- weatherdata::climate_full %>%
  filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%
  stretch() %>%
  filter(lubridate::year(date) == 2020) %>%
  tamp() %>%
  mutate(type = "climate")

river <- weatherdata::water %>%
  stretch() %>%
  select(date, Water_course_level) %>%
  rename(prcp = Water_course_level) %>% 
  tamp() %>%
  mutate(type = "river")
```

```{r matching-map, out.width="100%", fig.cap="asdfasd"}
knitr::include_graphics("figures/matching-map.png")
```

From the map, a few water gauges and weather stations are close to each other and we may be able to observe a spontaneous increase in the precipitation and water level from the measures, potential with a little delay in the river gauges. `match_sites()` matches one source of data with another source in both spatial and temporal dimension in a cubble.

<!-- Now we use `match_sites()` to first pair the weather stations with the river gauges spatially and then apply the temporal matching on `prcp`. We will construct the interval based on peaks in `climate` since we would expect a lag effect for precipitation to flow into the river and cause a raise in river level, hence `temporal_independent = climate`. We select the 30 highest peak from the series to construct the match by setting `temporal_n_highest = 30`. This is a tuning parameter and you can start with 10% of the points of one series (here we have daily data for a year, 10% is roughly 30 points). `temporal_min_match` filters out pairs don't have enough match and to return all the pairs, set `temporal_min_match` to `0`.`` -->

```{r echo = TRUE}
res <- match_sites(river, climate,
                   temporal_var_to_match = prcp,
                   temporal_independent = climate,  
                   temporal_n_highest = 30,
                   temporal_min_match = 15)
```

The output from temporal matching is also a cubble, with additional column `dist` and `group` produced from spatial matching and `n_match` from the number of peak matched temporally.

```{r}
res
```

Then plots can be made to inspect the matched pairs on the map or to view the co-movement of the two series:

```{r matching, out.width="100%", fig.cap="asdfasd"}
knitr::include_graphics("figures/matching.png")
```

With the chosen parameter, four pairs of stations stands out, which all locates in the middle Victoria and the concurrent increase of precipitation and water level can be observed.

### 2022-01-27 main.tex
improve aggregation plot
clean up water level example code and a draft on its example section

{\centering \includegraphics[width=1\linewidth,height=0.9\textheight]{figures/basic-agg}

@@

\caption[sdfasdf]{sdfasdf}\label{fig:basic-agg}
\end{figure}
\end{CodeChunk}

@@

The water level data comes from
\href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{Bureau
of Meteorology} and has a copy in \texttt{weatherdata}. Here we extract
the water course level and add a column annotate this data of type
\texttt{river}. For the rainfall data, we will still use the
\texttt{weatherdata::climate\_full}, filtering for Victorian stations in
2020 should be pretty familiar by now. Again, we first look at where
these stations are on the map first:

Now we use \texttt{match\_sites()} to first pair the weather stations
with the river gauges spatially and then apply the temporal matching on
\texttt{prcp}. We will construct the interval based on peaks in
\texttt{climate} since we would expect a lag effect for precipitation to
flow into the river and cause a raise in river level, hence
\texttt{temporal\_independent\ =\ climate}. We select the 30 highest
peak from the series to construct the match by setting
\texttt{temporal\_n\_highest\ =\ 30}. This is a tuning parameter and you
can start with 10\% of the points of one series (here we have daily data
for a year, 10\% is roughly 30 points). \texttt{temporal\_min\_match}
filters out pairs don't have enough match and to return all the pairs,
set \texttt{temporal\_min\_match} to \texttt{0}.``

The output from temporal matching is also a cubble, with additional
column \texttt{.dist} and \texttt{.group} inherent from spatial matching
and \texttt{n\_match} for the number of matched temporal peaks. Then you
can use this output to plot the location of match or to look at the
series:

@@

{\centering \includegraphics[width=1\linewidth,height=0.3\textheight]{figures/matching}

@@

>>>>

{\centering \includegraphics[width=1\linewidth]{figures/basic-agg}

@@

\caption[Aggregated precipitation in Australia]{Aggregated precipitation in Australia.}\label{fig:basic-agg}
\end{figure}
\end{CodeChunk}

@@

Water data is collected by the
\href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{Bureau
of Meteorology}. This includes \ldots{} water gauges, variables
collected. In particular, water level will interactive with the climate
variable precipitation as rains fall into the river will raise the water
level. Figure \ref{fig:matching-map} shows the available weather station
and water gauges in Victoria, Melbourne.

\begin{CodeChunk}
\begin{CodeInput}
R> climate <- weatherdata::climate_full %>%
+   filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%
+   stretch() %>%
+   filter(lubridate::year(date) == 2020) %>%
+   tamp() %>%
+   mutate(type = "climate")
R> 
R> river <- weatherdata::water %>%
+   stretch() %>%
+   select(date, Water_course_level) %>%
+   rename(prcp = Water_course_level) %>% 
+   tamp() %>%
+   mutate(type = "river")
\end{CodeInput}
\end{CodeChunk}

@@

{\centering \includegraphics[width=1\linewidth]{figures/matching-map}

\caption[asdfasd]{asdfasd}\label{fig:matching-map}
\end{figure}
\end{CodeChunk}

From the map, a few water gauges and weather stations are close to each
other and we may be able to observe a spontaneous increase in the
precipitation and water level from the measures, potential with a little
delay in the river gauges. \texttt{match\_sites()} matches one source of
data with another source in both spatial and temporal dimension in a
cubble.

\begin{CodeChunk}
\begin{CodeInput}
R> res <- match_sites(river, climate,
+                    temporal_var_to_match = prcp,
+                    temporal_independent = climate,  
+                    temporal_n_highest = 30,
+                    temporal_min_match = 15)
\end{CodeInput}
\end{CodeChunk}

The output from temporal matching is also a cubble, with additional
column \texttt{dist} and \texttt{group} produced from spatial matching
and \texttt{n\_match} from the number of peak matched temporally.

\begin{CodeChunk}
\begin{CodeOutput}
# cubble:   id [8]: nested form
# bbox:     [144.52, -37.73, 146.06, -36.55]
# temporal: date [date], prcp [dbl]
  id          name                lat  long ts         type   dist group n_match
  <chr>       <chr>             <dbl> <dbl> <list>     <chr> <dbl> <int>   <int>
1 405234      SEVEN CREEKS @ D~ -36.9  146. <tibble [~ river  6.15     5      21
2 ASN00082042 strathbogie       -36.8  146. <tibble [~ clim~  6.15     5      21
3 404207      HOLLAND CREEK @ ~ -36.6  146. <tibble [~ river  8.54    10      21
4 ASN00082170 benalla airport   -36.6  146. <tibble [~ clim~  8.54    10      21
5 230200      MARIBYRNONG RIVE~ -37.7  145. <tibble [~ river  6.17     6      19
# ... with 3 more rows
\end{CodeOutput}
\end{CodeChunk}

Then plots can be made to inspect the matched pairs on the map or to
view the co-movement of the two series:

{\centering \includegraphics[width=1\linewidth]{figures/matching}

@@

With the chosen parameter, four pairs of stations stands out, which all
locates in the middle Victoria and the concurrent increase of
precipitation and water level can be observed.
